---
title: "Project Writeup for Practical Machine Learning"
author: "Peggy Ong"
date: "Thursday, December 18, 2014"
output: html_document
---

Loading the libraries.
```{r,warning=FALSE,message=FALSE}
library(caret)
library(nnet)
```

 
Read Train and Score datasets. Get accelerometers data on the belt, forearm, 
arm, and dumbell of 6 participants.
```{r}
train_data <- read.csv('pml-training.csv',header = TRUE)
test_data <- read.csv('pml-testing.csv',header = TRUE)

train_data <- train_data[,c(grep(x = names(train_data),pattern = "^accel_"),160)]
sum(complete.cases(train_data)); dim(train_data)

test_data <- test_data[,c(grep(x = names(test_data),pattern = "^accel_"),160)]
sum(complete.cases(test_data)); dim(test_data)
```

 
Create data partitions to validate the models created. 70% of train data assign
to train partition and the remaining 30% to test partition.
```{r}
set.seed(888)
trainIndex <- createDataPartition(train_data$classe, p = .7,list = FALSE,
                                  times = 1)

train_partition <- train_data[trainIndex, ]
test_partition <- train_data[-trainIndex, ]

sapply(train_partition, sd)
sapply(test_partition, sd)
sapply(test_data, sd)
```
 
 Set the Formula 
```{r}
theTarget <- "classe"
theFormula <- as.formula(paste("as.factor(",theTarget, ") ~ . "))
trainTarget = train_partition[,which(names(train_partition)==theTarget)] 
testTarget  = test_partition[,which(names(test_partition)==theTarget)]

```
 
  
 As outcome is categorical with > 2 levels i.e. A, B, C, D or E, I use 
 multinomial logit model.
```{r}
multi_Model <- multinom(theFormula, data=train_partition)

train_pred <- predict(multi_Model, train_partition)
test_pred <- predict(multi_Model, test_partition)
```
 
 Display the confusion matrix results for train partition (multinomial).
```{r}
confusionMatrix(train_pred, trainTarget)
```
 
 Display the confusion matrix results for test partition (multinomial).
```{r}
confusionMatrix(test_pred, testTarget)
```
 
 Both Train and Test partitions under multinomial achieved only 50% accuracy. 
 That means in-sample error is very high at about 50%. 
 Using the test partition, we see that out-of-sample error  
 (model was not trained with this partition) is also high at 
 (1-0.5072)*100 = 49.28%.
 20 Test Cases are predicted with multinomial model achieved only 13/20 correct 
 in the submission.
```{r}
predict(multi_Model, test_data)
```
 
 Now, I try Generalized Boosted Regression Modeling (GBM) to see if better 
 results can be achieved.
```{r,message=FALSE, warning=FALSE}
gbm_Model <- train(theFormula, method="gbm", data=train_partition)

train_pred <- predict(gbm_Model, train_partition)
test_pred <- predict(gbm_Model, test_partition)
```

 Display the confusion matrix results for train partition (GBM).
```{r}
confusionMatrix(train_pred, trainTarget)
```
 
 Display the confusion matrix results for test partition (GBM).
```{r}
confusionMatrix(test_pred, testTarget)
```
 
 Both Train and Test partitions fitted with GBM algorithm achieved > 80% 
 accuracy! :) 
 That means in-sample error is much lower compared to Multinomial model.
 Out-of-sample error is also much lower at (1-0.8209)*100 = 17.91%.
 However, 20 Test Cases predicted with GBM model achieved only 14/20 correct.
```{r}
predict(gbm_Model, test_data)
```
